---
title: "14.310x: Data Analysis for Social Scientists - Homework 9"
output: html_notebook
---
\
\
Welcome to your ninth homework assignment! You will have about one week to work through the
assignment. We have provided this PDF copy of the assignment so that you can print and work through
the assignment offline. You can also go online directly to complete the assignment. If you choose to work
on the assignment using this PDF, please go back to the online platform to submit your answers based
on the output produced.\
Good luck :)!\

```{r}
# Preliminaries
rm(list = ls())
```

For the following questions, you will need the data set: [nlsw88.csv](https://prod-edxapp.edx-cdn.org/assets/courseware/v1/965fe0633ade16afe13e0190ab45a225/asset-v1:MITx+14.310x+3T2018+type@asset+block/nlsw88.csv). The data has information on labor market outcomes of a representative sample of women in the US. It contains the following variables: the logarithm of wage (lwage), total years of schooling (yrs_school), total experience in the labor markets (ttl_experience), and a dummy variable that indicates whether the woman is black or not. Since we are going to work with this data throughout this homework, please load it into R using the command read.csv\

```{r}
"Data file"
data_url <- "https://prod-edxapp.edx-cdn.org/assets/courseware/v1/965fe0633ade16afe13e0190ab45a225/asset-v1:MITx+14.310x+3T2018+type@asset+block/nlsw88.csv"
data_file <- "Data/nlsw88.csv"
if (!file.exists(data_file)) {
  download.file(data_url, destfile = data_file, method = "auto")
}
data_file

"Information on labor market outcomes of a representative sample of women in the US"
women <- read.csv(data_file)
women
summary(women)
```

As a first step, we are interested in estimating the following linear model:\
$$
log(wage_{i}) = \beta_{0} + \beta_{1} yrs\_school_{i} + \varepsilon_{i}
$$\
Estimate this equation by OLS using the command lm. Please go to the documentation in R to understand the syntax of the command. Based on your results, answer the following questions:\

**Question 1**\
According to this model, what is the estimate of $\beta_{1}$?\
Please round your answer to the third decimal point, i.e. if it is 0.12494, please round to 0.125 and if it is 0.1233, please round to 0.123\
```{r}
"Linear model for logarithm of wage (lwage) w.r.t total years of schooling (yrs_school)"
lwage_s.linear <- lm(lwage ~ yrs_school, data = women)
summary(lwage_s.linear)
```

**Question 2**\
What is the $90\%$ confidence interval (CI) of $\hat{\beta}_{1}$ according to this model?\
a. It is given by $[0.08174972,  0.1040900]$\
b. It is given by $[0.08736549, 0.09847428]$\
c. It is given by $[0.08442308, 0.1014167]$\
d. It is given by $[0.08579005, 0.1000497]$\
```{r}
"Confidence interval of Beta 1 estimate (associated with yrs_school)"
confint(lwage_s.linear, parm = "yrs_school", level = 0.9)
```

**Question 3**\
Assume that instead of having all the data, you just know that the covariance between the logarithm of the wage and the years of schooling is 0.6043267. What other information would you need to be able to find $\hat{\beta_{1}}$?\
a. The sample variance of the variable yrs_school\
b. The sample variance of the variable lwage\
c. The sample variance of the error term\
d. The sample covariance between the error term and yrs_school\
$$
\hat{\beta_{1}} = \frac{\frac{1}{n} \sum( x_{i} -\overline{x}) ( y_{i} -\overline{y})}{ \frac{1}{n} \sum 
  ( x_{i} -\overline{x})^{2} }
$$\
In the numerator is the expression for the covariance between x and y.\
In the doniminator is the expression for the variance of x.\

**Question 4**\
After running your code, what is the value you found for $\hat{\beta}_{0}?$\
Please round your answer to the third decimal point, i.e. if it is 0.12494, please round to 0.125 and if it is 0.1233, please round to 0.123\
```{r}
lwage_s.linear$coefficients
```

**Question 5**\
True or False: For any simple bivariate linear regression model, the predicted value when $x = \overline{x}$ is $\overline{y}$\
a. True\
b. False\
```{r}
"True"
```

**Question 6**\
After running your model, use the command residuals to calculate the residuals of the regression. Calculate the sum of the residuals. Should we be surprised that the sum is so close to zero?\
a. Yes\
b. No\
```{r}
"Residuals"
str(lwage_s.linear$residuals)

"Sum of residuals"
sum(lwage_s.linear$residuals)
```

Now, we are interested in estimating the following model:\
$$
\log(wage_{i}) = \beta_{0} + \beta_{1} black + \varepsilon_{i}
$$\

**Question 7**\
Researcher A says that this model is not correctly specified. Researcher A suggests that the correct model should estimate the following equation (where $other \; race$ is a dummy variable equal to 1 when the person is not black):\
$$
log(wage_{i}) = \beta_{0} + \beta_{1} black + \beta_{2} other \;  race + \varepsilon_{i}
$$\
Researcher B claims that Researcher A is wrong, and that in this second model, it is not possible to separately identify $\beta_{0},\; \beta_{1},\; \text{and } \beta_{2}.$ Who do you agree with?\
a. Researcher A\
b. Researcher B\
```{r}
"Researcher B"
```

**Question 8**\
Assume that you don't have all the data. However, you know that the sample mean of the log wage for women who are not black is $\overline{y}_{other}$, and the sample mean of the log wage for black women is $\overline{y}_{black}.$ What are the values of $\hat{\beta}_{0}$ and $\hat{\beta}_{1}$ if we run this model using OLS?\
a. We have that $\hat{\beta}_{0}=\overline{y}_{black}$ and $\hat{\beta}_{1}=\overline{y}_{other}-\overline{y}_{other}$\
b. We have that $\hat{\beta}_{0}=\overline{y}_{black}$ and $\hat{\beta}_{1}=\overline{y}_{black}-\overline{y}_{other}$\
c.We have that $\hat{\beta}_{0}=\overline{y}_{other}$ and $\hat{\beta}_{1}=\overline{y}_{black}-\overline{y}_{other}$\
d. We have that $\hat{\beta}_{0}=\overline{y}_{other}$ and $\hat{\beta}_{1}=\overline{y}_{other}-\overline{y}_{other}$\
```{r}
"c"
```

**Question 9**\
Now, estimate this model by yourself using both the sample means approach and the regression approach with the command lm. (You should get the same results!)\
For the following answers, please round to the third decimal place, i.e. if the solution is 0.23412, please round to 0.234, and if it is 0.23498, please round to 0.235.\
What value did you find for $\hat{\beta}_{0}$?\
```{r}
"Linear model for logarithm of wage (lwage) w.r.t race (black)"
lwage_b.linear <- lm(lwage ~ black, data = women)
summary_lwage_b <- summary(lwage_b.linear)
summary_lwage_b

"Means of logarithm of wage (lwage) by race (black)"
lwage_means_b <- ddply(women, .variables = .(black), .fun = summarise, lwage_mean = mean(lwage))
lwage_means_b

"Coefficient Beta 0 (using regression approach)"
lwage_b.linear$coefficients["(Intercept)"]

"Coefficient Beta 0 (using sample means approach)"
filter(lwage_means_b, black == 0)$lwage_mean
```
What value did you find for $\hat{\beta}_{1}$?\
```{r}
"Coefficient Beta 1 (on black) (using regression approach)"
lwage_b.linear$coefficients["black"]

"Coefficient Beta 1 (on black) (using sample means approach)"
filter(lwage_means_b, black == 1)$lwage_mean - filter(lwage_means_b, black == 0)$lwage_mean
```

**Question 10**\
A critic is claiming that this doesn't prove that there are differences in the wage of black women and women of other races. You decide to conduct a test on the parameter $\beta_{1},$ where the null hypothesis is $\beta_{1} =0.$ What is the value of the statistic of this test?\
Please round to the third decimal place, i.e. if the solution is 0.23412, please round to 0.234, and if it is 0.23498, please round to 0.235.\
```{r}
"All coefficients (matrix)"
summary_lwage_b$coefficients

"t-statistic of sample under null 'Beta 1 (on black) = 0'"
summary_lwage_b$coefficients["black", "t value"]
```

**Question 11**\
Would you reject this null hypothesis using a $99\%$ level of confidence?\
a. Yes\
b. No\
```{r}
"All coefficients (matrix)"
summary_lwage_b$coefficients

"Significance level of sample under null 'Beta 1 (on black) = 0' (Probability of observing t > |t-statistic|)"
signif_level_obs <- summary_lwage_b$coefficients["black", "Pr(>|t|)"]
signif_level_obs

"Significance level required by 99% confidence level"
signif_level_req <- 1 - 0.99
signif_level_req

"Probability that observation on sample is due to chance, under null 'Beta 1 (on black) = 0', is low (<= required)?"
signif_level_obs <= signif_level_req
```

Labor economists have estimated Mincer equations that include not only total years of schooling, but also total experience as explanatory variables of the wage. Assume now that you want to estimate the following model:\
$$
log(wage_{i}) = \beta_{0} + \beta_{1} yrs\_school_{i} + \beta_{2} total \; experience + \varepsilon_{i}
$$\

**Question 12**\
If you run this model in R, what would be the value of the $R^{2}$?\
Please round your answer to the third decimal place, i.e. if your answer is 0.7283, please round to 0.728 and if it is 0.7289, round to 0.729.\
```{r}
"Linear model for logarithm of wage (lwage) w.r.t total years of schooling (yrs_school), total experience in the labor markets (ttl_exp)"
lwage_s_e.linear <- lm(lwage ~ yrs_school + ttl_exp, data = women)
summary_lwage_s_e <- summary(lwage_se.linear)
summary_lwage_s_e

"R-squared"
summary_lwage_s_e$r.squared
```

Some young folks are claiming that they prefer to drop out from school since each additional year of schooling changes the log of the wage in the same amount as one half year of experience. A group of parents are really worried. They ask you to conduct a formal test over this sample.\

**Question 13**\
What would be the null hypothesis of this test?\
a. The null hypothesis of this test is: $\beta_{1} = 2 \beta_{2}$\
b. The null hypothesis of this test is: $\beta_{1} =  \beta_{2} + \beta_{1}$\
c. The null hypothesis of this test is: $\beta_{1} + \beta_{2} =  \beta_{2}$\
d. The null hypothesis of this test is: $2 \beta_{1} =  \beta_{2}$\
```{r}
"d"
"Null hypothesis:"
"(Beta 1) * (1 additional yrs_school) = (Beta 2) * (0.5 additional ttl_exp)"
"(Beta 1) * (2 additional yrs_school) = (Beta 2) * (1 additional ttl_exp)"
"2(Beta 1) = (Beta 2)"
```

**Question 14**\
Which of the following would correspond to the restricted model under this null hypothesis? (Select all that apply)\
a. The model $log(wage_{i}) = \beta_{0} +   \beta_{2} (  yrs\_school_{i}  + 2 total \; experience_{i} ) + \varepsilon_{i}$\
b. The model $log(wage_{i}) = \beta_{0} +   \beta_{1} ( \frac{1}{2} yrs\_school_{i} +  total \; experience_{i} ) + \varepsilon_{i}$\
c. The model $log(wage_{i}) = \beta_{0} +   \beta_{1} (   yrs\_school_{i}  + 2 total \; experience_{i} ) + \varepsilon_{i}$\
d. The model $log(wage_{i}) = \beta_{0} + (  \beta_{1} + 2\beta_{2} ) yrs\_school_{i}  + \varepsilon_{i}$\
e. The model $log(wage_{i}) = \beta_{0} + (   2 \beta_{1} + \beta_{2} ) yrs\_school_{i}   + \varepsilon_{i}$\
f. The model $log(wage_{i}) = \beta_{0} +   \beta_{2} \left( \frac{1}{2} yrs\_school_{i} +  total \; experience_{i} \right) + \varepsilon_{i}$\
```{r}
"c, f"
"Possible restricted models, under null '2(Beta 1) = (Beta 2)', can be obtained by plugging null hypothesis (2(Beta 1) = (Beta 2)) into original model"
```

**Question 15**\
Estimate the restricted model in R. What is the value that you obtain for $\hat{\beta}_{1}$ in the restricted model? For the restricted model, use the first correct model in the list from question 14, i.e. if both (a) and (d) are correct, use (a) or if both (b) and (c) are correct, use (b).\
Please round your answer to the fourth decimal place, i.e. if your answer is 0.78244, please round to 0.7824, and if it is 0.78247, please round to 0.7825.\
Restricted model under null $2 \beta_{1} =  \beta_{2}$:\
$$
log(wage_{i}) = \beta_{0} +   \beta_{1} (   yrs\_school_{i}  + 2 total \; experience_{i} ) + \varepsilon_{i}
$$\
```{r}
"Data for restricted model under null '2(Beta 1) = (Beta 2)' (calculated values of (total years of schooling + 2 * total experience in the labor markets))"
women$s_add_2e <- women$yrs_school + women$ttl_exp
women
summary(women)

lwage_.linear <- lm(lwage ~ yrs_school, data = women)
```









