---
title: "14.310x: Data Analysis for Social Scientists - Homework 9"
output: html_notebook
---


Welcome to your ninth homework assignment! You will have about one week to work through the
assignment. We have provided this PDF copy of the assignment so that you can print and work through
the assignment offline. You can also go online directly to complete the assignment. If you choose to work
on the assignment using this PDF, please go back to the online platform to submit your answers based
on the output produced.
Good luck :)!

```{r}
# Preliminaries
rm(list = ls())
```

For the following questions, you will need the data set: [nlsw88.csv](https://prod-edxapp.edx-cdn.org/assets/courseware/v1/965fe0633ade16afe13e0190ab45a225/asset-v1:MITx+14.310x+3T2018+type@asset+block/nlsw88.csv). The data has information on labor market outcomes of a representative sample of women in the US. It contains the following variables: the logarithm of wage (lwage), total years of schooling (yrs_school), total experience in the labor markets (ttl_experience), and a dummy variable that indicates whether the woman is black or not. Since we are going to work with this data throughout this homework, please load it into R using the command read.csv

```{r}
"Data file"
data_url <- "https://prod-edxapp.edx-cdn.org/assets/courseware/v1/965fe0633ade16afe13e0190ab45a225/asset-v1:MITx+14.310x+3T2018+type@asset+block/nlsw88.csv"
data_file <- "Data/nlsw88.csv"
if (!file.exists(data_file)) {
  download.file(data_url, destfile = data_file, method = "auto")
}
data_file

"Information on labor market outcomes of a representative sample of women in the US"
women <- read.csv(data_file)
women
summary(women)
```

As a first step, we are interested in estimating the following linear model:
$$
log(wage_{i}) = \beta_{0} + \beta_{1} yrs\_school_{i} + \varepsilon_{i}
$$
Estimate this equation by OLS using the command lm. Please go to the documentation in R to understand the syntax of the command. Based on your results, answer the following questions:

**Question 1**
According to this model, what is the estimate of $$\beta_{1}$$?
Please round your answer to the third decimal point, i.e. if it is 0.12494, please round to 0.125 and if it is 0.1233, please round to 0.123
```{r}
"Regress logarithm of wage (lwage) from total years of schooling (yrs_school)"
lwage_s.linear <- lm(lwage ~ yrs_school, data = women)
lwage_s.linear
summary(lwage_s.linear)
```

**Question 2**
What is the $$90\%$$ confidence interval (CI) of $$\hat{\beta}_{1}$$ according to this model?
a. It is given by $$[0.08174972,  0.1040900]$$
b. It is given by $$[0.08736549, 0.09847428]$$
c. It is given by $$[0.08442308, 0.1014167]$$
d. It is given by $$[0.08579005, 0.1000497]$$
```{r}
"Confidence interval of Beta 1 estimate (associated with yrs_school)"
confint(lwage_s.linear, parm = "yrs_school", level = 0.9)
```

**Question 3**
Assume that instead of having all the data, you just know that the covariance between the logarithm of the wage and the years of schooling is 0.6043267. What other information would you need to be able to find $$\hat{\beta_{1}}$$?
a. The sample variance of the variable yrs_school
b. The sample variance of the variable lwage
c. The sample variance of the error term
d. The sample covariance between the error term and yrs_school
```{r}

```

**Question 4**
After running your code, what is the value you found for $$\hat{\beta}_{0}?$$
Please round your answer to the third decimal point, i.e. if it is 0.12494, please round to 0.125 and if it is 0.1233, please round to 0.123
```{r}
lwage_s.linear$coefficients
```

**Question 5**
True or False: For any simple bivariate linear regression model, the predicted value when $$x = \overline{x}$$ is $$\overline{y}$$
a. True
b. False

**Question 6**
After running your model, use the command residuals to calculate the residuals of the regression. Calculate the sum of the residuals. Should we be surprised that the sum is so close to zero?
a. Yes
b. No






